EV Predictive Modeling

### 2024-08-06 code/analysis-code/statistical_analysis.qmd
add re-render for organization plus minor edits

---
title: "Statistical Analysis"
date: today
output: html_document
execute: 
  error: False
  warning: False
---

@@

rand_for_feature_importance_list = xgboost_feature_importance['Feature'].head(10).tolist()
print(rand_for_feature_importance_list)

# Plotting function
def plot_feature_importance(importances, model_name):
    plt.figure(figsize=(15, 10))
    plt.barh(importances['Feature'], importances['Importance'], color='skyblue')
    plt.xlabel('Importance')
    plt.ylabel('Feature')
    plt.title(f'Feature Importance - {model_name}')
    plt.gca().invert_yaxis()  # Invert y-axis to have the most important feature at the top
    plt.grid(True)
    plt.tight_layout()
    plt.savefig(f'../../results/output/variable_importance/{model_name}.png')

@@

# Display the comparison results
print(model_comparison_metrics_df)
model_comparison_metrics_df.to_csv('../../results/output/model_comparison.csv', index=False)

>>>>

---
title: "Statistical Analysis"
output: html_document
execute: 
  error: False
  warning: False
---

@@

# Plotting function
def plot_feature_importance(importances, model_name):
    plt.figure(figsize=(10, 5))
    plt.barh(importances['Feature'], importances['Importance'], color='skyblue')
    plt.xlabel('Importance')
    plt.ylabel('Feature')
    plt.title(f'Feature Importance - {model_name}')
    plt.gca().invert_yaxis()  # Invert y-axis to have the most important feature at the top
    plt.grid(True)
    plt.tight_layout()
    plt.savefig(f'../../results/output/variable_importance/{model_name}.png')
    plt.show()

@@

# Display the comparison results
model_comparison_metrics_df.to_csv('../../results/output/model_comparison.csv', index=False)
```

### Model Comparison Results
```{python}
print(model_comparison_metrics_df)
```

### 2024-08-06 code/eda-code/eda-ev-tax.qmd
add re-render for organization plus minor edits

---
title: "Log-Transform Predictors"
date: today
output: html_document
execute: 
  error: False
  warning: False
---
**Electric Vehicle Data Analysis**

# Introduction
Apply log transformation to our feature

# Setup

```{python}
import pandas as pd
from sklearn.model_selection import train_test_split
import re

@@

# Splitting the data into training and testing sets
X_train_grouped, X_test_grouped, y_train_grouped, y_test_grouped = train_test_split(
    X_grouped, y_grouped, test_size=0.2, random_state=42)

merged_data.to_csv('test.csv', index=False)
```
<!-- 
```{python}
import matplotlib.pyplot as plt

# Scatter plots to understand relationship between predictors and independent variable
for i, feature in enumerate(features_to_use):
    plt.figure(figsize=(10,5))
    plt.scatter(X_grouped[feature], y_grouped, c="lightblue", edgecolors="blue")
    plt.title(f'{feature} vs. Vehicle Count Relationship')
    plt.xlabel(feature)
    plt.ylabel('Vehicle Count')
    # plt.grid(True)
    plt.savefig(f'../../results/figures/scatterplots/{feature}_scatterplot.png')

```

```{python}
import numpy as np
import pandas as pd
from sklearn.impute import SimpleImputer
import matplotlib.pyplot as plt
import seaborn as sns

@@

    for feature in data.columns:
        fig, ax = plt.subplots(1, 2, figsize=(10,5))
        ax1 = plt.subplot(2,1,1)
        ax1 = sns.histplot(data[feature], kde=True, bins=50)
        ax2 = plt.subplot(2,1,2)
        ax2 = sns.histplot(data1[feature], kde=True, bins=50)
        ax1.title.set_text(f'Original vs. Log-Transformed: {feature}')
        # ax2.title.set_text(f'Log-Transformed Histogram of {feature}')
        plt.xlabel(feature)
        plt.ylabel('Frequency')
        plt.grid(True)
        plt.savefig(f'../../results/figures/histograms/{feature}_histogram.png')

@@

```{python}
# Plot boxplot for each feature in the transformed dataset
def plot_transformed_feature_boxplots(data, data1):

    for feature in data.columns:
        fig, ax = plt.subplots(1, 2, figsize=(10,5))
        ax1 = plt.subplot(2,1,1)
        ax1 = sns.boxplot(data[feature])
        ax2 = plt.subplot(2,1,2)
        ax2 = sns.boxplot(data1[feature])
        ax1.title.set_text(f'Original vs. Log-Transformed Boxplot of {feature}')
        # ax2.title.set_text(f'Log-Transformed Boxplot of {feature}')
        plt.xlabel(feature)
        plt.ylabel('Frequency')
        plt.grid(True)
        plt.tight_layout()
        plt.savefig(f'../../results/figures/boxplots/{feature}_boxplot.png')

# Plot the boxplots for the log-transformed features
plot_transformed_feature_boxplots(X_grouped, X_transformed)
``` -->

>>>>

---
title: "Log-Transform Predictors"
output: html_document
execute: 
  error: False
  warning: False
---
**Electric Vehicle Data Analysis**

## Introduction
Create scatterplots to see the relationship between the independent variable and predictors. Perform a log transformation on features to normalize and compare original to transformed distribution.

## Import Libraries & Data, Create Train/Test Datasets
```{python}
import pandas as pd
from sklearn.model_selection import train_test_split
import re

@@

# Splitting the data into training and testing sets
X_train_grouped, X_test_grouped, y_train_grouped, y_test_grouped = train_test_split(
    X_grouped, y_grouped, test_size=0.2, random_state=42)
```

## Scatterplot Relationship - Predictor vs. Independent Variable
```{python}
import matplotlib.pyplot as plt

# Scatter plots to understand relationship between predictors and independent variable
for i, feature in enumerate(features_to_use):
    plt.figure(figsize=(10,5))
    plt.scatter(X_grouped[feature], y_grouped, c="lightblue", edgecolors="blue")
    plt.title(f'{feature} vs. Vehicle Count Relationship')
    plt.xlabel(feature)
    plt.ylabel('Vehicle Count')
    plt.savefig(f'../../results/figures/scatterplots/{feature}_scatterplot.png')
    plt.show()
```

## Histogram Distributions
```{python}
import numpy as np
import pandas as pd
from sklearn.impute import SimpleImputer
import matplotlib.pyplot as plt
import seaborn as sns

@@

    for feature in data.columns:
        fig, ax = plt.subplots(1, 2, figsize=(10,5))
        ax1 = plt.subplot(2,1,1)
        ax1 = sns.histplot(data[feature], kde=True, bins=50)
        ax2 = plt.subplot(2,1,2)
        ax2 = sns.histplot(data1[feature], kde=True, bins=50)
        ax1.title.set_text(f'Original vs. Log-Transformed: {feature}')
        plt.xlabel(feature)
        plt.ylabel('Frequency')
        plt.grid(True)
        plt.tight_layout()
        plt.savefig(f'../../results/figures/histograms/{feature}_histogram.png')
        plt.show()

@@

## Boxplot Distributions
```{python}
# Plot boxplot for each feature in the transformed dataset
def plot_transformed_feature_boxplots(data, data1):

    for feature in data.columns:
        fig, ax = plt.subplots(1, 2, figsize=(10,5))
        ax1 = plt.subplot(2,1,1)
        ax1 = sns.boxplot(data[feature])
        ax2 = plt.subplot(2,1,2)
        ax2 = sns.boxplot(data1[feature])
        ax1.title.set_text(f'Original vs. Log-Transformed Boxplot of {feature}')
        plt.xlabel(feature)
        plt.ylabel('Frequency')
        plt.grid(True)
        plt.tight_layout()
        plt.savefig(f'../../results/figures/boxplots/{feature}_boxplot.png')
        plt.show()

# Plot the boxplots for the log-transformed features
plot_transformed_feature_boxplots(X_grouped, X_transformed)
```

### 2024-08-06 index.qmd
add re-render for organization plus minor edits

---
title: "Using IRS Data to Predict Electric Vehicle Count"
author: Sri Lakshmi Chundru, Collin Real, Joaquin Ramirez, Seth Harris
bibliography: assets/dataanalysis-references.bib
csl: assets/apa.csl
---

@@

1. Self-Employed Health Insurance Deduction Amount
This feature represents the amount self-employed individuals can deduct for health insurance premiums. It is significant as it indicates the presence and financial activities of self-employed individuals within a zip code. Self-employed individuals often have variable incomes and unique financial situations, which may influence their ability to purchase and maintain vehicles. This deduction could correlate with areas where self-employment is prevalent, possibly reflecting regions with more entrepreneurial activity and, potentially, higher disposable incomes.

@@

3. Home Mortgage from Personal Seller Amount
This feature involves transactions where the home mortgage is obtained directly from a personal seller rather than a traditional financial institution. It indicates non-standard financial arrangements, possibly reflecting unique housing markets or economic conditions in certain areas. The relevance of this feature could be tied to the accessibility and flexibility in financial transactions within these regions, potentially influencing overall economic activity, including vehicle purchases.

4. Salaries and Wages Amount
A direct indicator of income, this feature represents the total salaries and wages reported within a zip code. It is a critical predictor as higher salaries and wages generally correlate with greater financial capacity to purchase and maintain vehicles. This feature strongly influences vehicle ownership levels, including the adoption of EVs, which may have higher upfront costs but offer long-term savings.

5. Home Mortgage Interest Paid Amount
The amount of interest paid on home mortgages is indicative of both property ownership and the financial burden of maintaining such properties. Higher mortgage interest payments suggest higher home values and, by extension, potentially higher wealth levels. This wealth is a significant factor in vehicle ownership, as individuals with more expensive homes often have higher disposable incomes, enabling them to invest in additional vehicles or more luxurious options, including EVs.

6. Self-Employment Tax Amount
This feature captures the amount of self-employment tax paid, which directly correlates with self-employment income. Higher self-employment tax payments indicate successful businesses or high-earning self-employed individuals. The economic health of these individuals likely contributes to their ability to afford vehicles, making this a significant predictor of vehicle counts.

@@

8. Business or Professional Net Income (Less Loss) Amount
This feature represents the net income from business or professional activities, after accounting for losses. It is a measure of the economic success and profitability of businesses within a zip code. Higher net incomes suggest prosperous business environments, which can lead to increased vehicle ownership among business owners and employees, reflecting economic prosperity in these areas.

9. Real Estate Taxes Amount
Real estate taxes are directly linked to property values and indirectly to wealth. Higher real estate taxes generally indicate more expensive properties, suggesting higher overall wealth in an area. Wealthier populations have greater access to credit and financing options, which can increase the likelihood of purchasing vehicles, including newer and more expensive models like EVs.

10. Net Premium Tax Credit Amount
The net premium tax credit amount provides a tax credit under the Affordable Care Act, generally aimed at reducing the cost of health insurance for lower to middle-income families. While it may seem counterintuitive, this feature could correlate with regions where financial assistance is necessary, reflecting economic diversity. In these areas, the adoption of vehicles, including EVs, may be influenced by financial incentives or subsidies aimed at promoting energy-efficient technologies.

>>>>

---
title: "EV Predictive Modeling"
subtitle: "IRS Tax Data - Zip Code"
author: Sri Lakshmi Chundru, Collin Real, Joaquin Ramirez, Seth Harris
bibliography: assets/dataanalysis-references.bib
csl: assets/apa.csl
---

@@

1. Self-Employed Health Insurance Deduction Amount:
This feature represents the amount self-employed individuals can deduct for health insurance premiums. It is significant as it indicates the presence and financial activities of self-employed individuals within a zip code. Self-employed individuals often have variable incomes and unique financial situations, which may influence their ability to purchase and maintain vehicles. This deduction could correlate with areas where self-employment is prevalent, possibly reflecting regions with more entrepreneurial activity and, potentially, higher disposable incomes.

@@

3. Home Mortgage from Personal Seller Amount:
This feature involves transactions where the home mortgage is obtained directly from a personal seller rather than a traditional financial institution. It indicates non-standard financial arrangements, possibly reflecting unique housing markets or economic conditions in certain areas. The relevance of this feature could be tied to the accessibility and flexibility in financial transactions within these regions, potentially influencing overall economic activity, including vehicle purchases.

4. Salaries and Wages Amount:
A direct indicator of income, this feature represents the total salaries and wages reported within a zip code. It is a critical predictor as higher salaries and wages generally correlate with greater financial capacity to purchase and maintain vehicles. This feature strongly influences vehicle ownership levels, including the adoption of EVs, which may have higher upfront costs but offer long-term savings.

5. Home Mortgage Interest Paid Amount:
The amount of interest paid on home mortgages is indicative of both property ownership and the financial burden of maintaining such properties. Higher mortgage interest payments suggest higher home values and, by extension, potentially higher wealth levels. This wealth is a significant factor in vehicle ownership, as individuals with more expensive homes often have higher disposable incomes, enabling them to invest in additional vehicles or more luxurious options, including EVs.

6. Self-Employment Tax Amount:
This feature captures the amount of self-employment tax paid, which directly correlates with self-employment income. Higher self-employment tax payments indicate successful businesses or high-earning self-employed individuals. The economic health of these individuals likely contributes to their ability to afford vehicles, making this a significant predictor of vehicle counts.

@@

8. Business or Professional Net Income (Less Loss) Amount:
This feature represents the net income from business or professional activities, after accounting for losses. It is a measure of the economic success and profitability of businesses within a zip code. Higher net incomes suggest prosperous business environments, which can lead to increased vehicle ownership among business owners and employees, reflecting economic prosperity in these areas.

9. Real Estate Taxes Amount:
Real estate taxes are directly linked to property values and indirectly to wealth. Higher real estate taxes generally indicate more expensive properties, suggesting higher overall wealth in an area. Wealthier populations have greater access to credit and financing options, which can increase the likelihood of purchasing vehicles, including newer and more expensive models like EVs.

10. Net Premium Tax Credit Amount:
The net premium tax credit amount provides a tax credit under the Affordable Care Act, generally aimed at reducing the cost of health insurance for lower to middle-income families. While it may seem counterintuitive, this feature could correlate with regions where financial assistance is necessary, reflecting economic diversity. In these areas, the adoption of vehicles, including EVs, may be influenced by financial incentives or subsidies aimed at promoting energy-efficient technologies.

### 2024-08-06 products/manuscript/Manuscript.qmd
add re-render for organization plus minor edits

---
title: "Using IRS Data to Predict Electric Vehicle Count"
author: Sri Lakshmi Chundru, Collin Real, Joaquin Ramirez, Seth Harris
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/apa.csl
---

>>>>

---
title: "EV Predictive Modeling"
subtitle: "IRS Tax Data - Zip Code"
author: Sri Lakshmi Chundru, Collin Real, Joaquin Ramirez, Seth Harris
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/apa.csl
---
